---
title: "GLG 803 Homework 1: Reading and Writing Data into Dataframes"
output: html_notebook
author: Jake Roush
---


## 1) Reading in Odyssey datafile and cleaning up
```{r eval=TRUE}
# Reading in .csv  Odyssey data file
# Skipping first 6 lines - only site information
raw_data <- read.csv("S:/Users/roushjac/Misc/Class/GLG 803/Spring-2018/Homework/Homework_1/38085_001.PRN", skip = 6, as.is = c(1:7))
raw_data$X.2 <- NULL
# Renaming columns to legible names
var_names <- c("Scan_Num","Date_Time","Temp_Raw","Temp_Calib","Pres_Raw","Pres_Calib")
colnames(raw_data) <- var_names
# Deleting extra row
raw_data <- raw_data[-c(1), ]
```

## 2) Converting to appropriate dataypes
```{r eval=TRUE}
# Converting excel datetime to seconds before converting to R datetime - don't know why it needs seconds but it works???
raw_data$Date_Time <- as.POSIXct(((raw_data$Date_Time)*(60*60*24)), origin="1899-12-30", tz='GMT')
raw_data$Temp_Raw <- as.numeric(raw_data$Temp_Raw)
raw_data$Temp_Calib <- as.numeric(raw_data$Temp_Calib)
raw_data$Pres_Raw <- as.numeric(raw_data$Pres_Raw)
raw_data$Pres_Calib <- as.numeric(raw_data$Pres_Calib)
```

## 3) Head of the Odyssey dataframe

```{r eval=TRUE}
head(raw_data)
```

## 4, 5) Converting calibrated pressure from ft H2O to mm H2O, and temperature from C to F

```{r eval=TRUE}
FtToMm <- 304.8
raw_data$Pres_Ft <- raw_data$Pres_Calib / FtToMm
raw_data$Temp_F <- raw_data$Temp_Calib * (9/5) * 32
```

## 6) Opening native-format reference file

```{r eval=TRUE}
ref_data <- load("S:/Users/roushjac/Misc/Class/GLG 803/Spring-2018/Homework/Homework_1/reference_data.RData")
```

## 7) Sanity checks:

Comparing lengths of my file with reference file:
```{r eval=TRUE}
nrow(raw_data) # my data
nrow(dfReference) # reference data
```

Comparing mean values of pressure fields in ft H2O:
```{r eval=TRUE}
mean(raw_data$Pres_Ft) # my data
mean(dfReference$Pressure_ft) # reference data
```

Comparing first dates of both dataframes:
```{r eval=TRUE}
raw_data$Date_Time[1] # my data
dfReference$Datetime[1] # reference data
```

## 8) Exporting final .csv
```{r eval=TRUE}
write.csv(raw_data, file = "S:/Users/roushjac/Misc/Class/GLG 803/Spring-2018/Homework/Homework_1/Roush_HW1/gauge_data.csv")
```
